Component-based Software Development Framework \\for Embedded IoT Devices（仮）

### 2017-11-08 ipsj/paper.tex
Add TINET+TECS

@@

\subsubsection{Component Design of TLSF}

>>>>

\subsection{TINET+TECS}
\label{sec:TINET+TECS}

\subsubsection{TINET}

TINET is a compact TCP/IP protocol stack for embedded systems based on the ITRON\footnote{ITRON is a real-time operating system (RTOS) developed by the TRON project.} TCP/IP API Specification \cite{url:ITRON_TCP/IP_API_Spec}, developed by the TOPPERS Project \cite{url:TOPPERS}.
TINET has been released as an open-source tool.

To satisfy restrictions for embedded systems in terms of, for example, memory capacity, size, and power consumption, TINET supports the following functions:

\begin{itemize}
    \item minimum copy frequency,
    \item elimination of dynamic memory control,
    \item asynchronous interfacing,
    \item error detailing per API.
\end{itemize}

{\bf Overview:}
TINET runs as middleware on TOPPERS/ASP3 \cite{par:ASP3} \cite{url:ASP3}, a real-time kernel based on $\mu$ITRON \cite{par:microITRON}.
As it is compatible with TOPPERS RTOS, TINET also supports other RTOSs such as TOPPERS/ASP and TOPPERS/JSP.

Fig.\ref{fig:TINETHierarchyDiagram} shows the hierarchy diagram of TINET and TOPPERS/ASP3.
Users transmit and receive data using a Communication End Point (CEP), an interface that functions like a socket.
In the transmission process, headers are attached to the data body passed to the CEP at each protocol layer before the data are transmitted from the network device.
In the reception process, the headers of the data bodies received by the network device are analyzed at each protocol layer, and the data are then passed to the CEP.

A TCP reception point called the REP stands by to receive connection requests from the partner side.
The REP has an IP address ({\it myaddr}) and a port number ({\it myportno}) as attributes and performs functions such {\it bind()} and {\it listen()}.

In TINET, the amount of data copying between each protocol layers is minimized.
In standard computing systems, the TCP/IP protocol stack has large overheads in terms of execution time and memory consumption because the data are copied at each protocol layer.
To solve this problem, TINET does pass the pointer of the data buffer between each protocol layer instead of performing data copying.

\begin{figure}[t]
    \centering
    \includegraphics[width=7.0cm,clip]{figure/TINETHierarchyDiagram.pdf}
    \caption{TINET and TOPPERS/ASP3 hierarchy diagrams}
    \label{fig:TINETHierarchyDiagram}
\end{figure}

\subsubsection{Component Design of TINET+TECS}

TINET+TECS, the proposed componentized TCP/IP protocol stack, comprises a number of some TECS components.
This section describes the components of the TINET+TECS framework with the aid of component diagrams.

\subsubsection*{Components of a protocol stack}

The components of a TINET+TECS protocol stack are shown in Fig.\ref{fig:ComponentProtocolStack}.
Note that some small particle components, such as a kernel object, data queues, and semaphores, are omitted to simplify the component diagram.
In TINET+TECS, the components are divided for each protocol, and functionalities such as input and output functions are defined as respective components.
By using such small grain components, software visibility is improved.
The components of each protocol are described in the following.

\begin{figure}[t]
    \centering
    \includegraphics[width=8.0cm,clip]{figure/ComponentProtocolStack.pdf}
    \caption{Component diagram of a protocol stack}
    \label{fig:ComponentProtocolStack}
\end{figure}

{\bf Application layer:}
An application in TINET+TECS is implemented as a component such as tApplication.
Software with TINET uses ITRON TCP/IP API \cite{url:ITRON_TCP/IP_API_Spec} such as {\it tcp\_snd\_dat} and {\it tcp\_rcv\_dat}.
In TINET+TECS, the application component calls TECS functions such as {\it cTCPAPI\_sendData} and {\it cTCPAPI\_receiveData}.
Moreover, in TINET+TECS supporting a TECS adapter, an existing application with TINET can run on the TINET+TECS framework without transporting, and therefore, software can be developed either using existing methods or as TECS components.

{\bf Transport layer:}
tTCPCEP (tUDPCEP) and tREP are, respectively, CEP and REP components TCP (UDP).
For example, a server program supporting multiple clients can be developed by preparing multiple tTCPCEP components.
tTCPInput (tUDPInput) and tTCPOutput (tUDPOutput) are components for performing, respectively, receiving and sending processing in the transport layer.

{\bf Network layer:}
The tIPv4Input and tIPv4Output components perform, respectively, the receiving and sending processing in the network layer.
The tIPv4Functions component performs functions such as checksum, the tICMP component is used for the Internet Control Message Protocol (ICMP), and the tIPv4RoutingTable component operates a routing table.

{\bf Data link layer:}
tEthernetInputTaskBody and tEthernetOutputTaskBody (tEthernetOutput) are components for performing, respectively, receiving and sending processing in the data link layer.
The tArp component is for implementing the Address Resolution Protocol (ARP).

{\bf Physical layer:}
The tNetworkInterfaceContoroller component implements a network device driver.
Software can be run on other devices by replacing the component because only the component depends on the target device.

To utilize the protocol stack in the same manner in the original TINET, communication object components such as tTCPCEP, tUDPCEP, and tREP are defined as an interface between TINET+TECS and an application.
The communication object component corresponds to a CEP or REP of the original TINET.
Application developers can utilize the TINET+TECS functionalities by generating and combining as many components as necessary.

The protocol stack of TINET+TECS supports the coexistence of multiple protocols.
Though its use of IPv6 and Point-to-Point Protocol (PPP) components, TINET+TECS can make IPv4 and IPv6 coexist and support PPP without modification of component implementation.

\subsubsection*{Memory allocator component}

The original TINET eliminates dynamic memory control to meet the severe memory restrictions of embedded systems.
A memory area for sending/receiving data in the protocol stack is allocated and released within a predetermined area.
The memory allocator component allows for elimination of dynamic memory control in TINET+TECS by providing a requested memory area from the statically allocated memory area.

The memory allocator component connects to as many tFixedSizeMemoryPool as required, as shown in Fig.\ref{fig:tMemoryAllocator}.
tFixedSizeMemoryPool is a componentized kernel object of TOPPERS/ASP3 for allocating and release memory areas of a requested size. 
tFixedSizeMemoryPool components of various sizes are prepared, and an appropriate memory area can be allocated according to the used data size.
On the other hand, all components that need to allocate or deallocate memory, e.g., tTCPInput and tEthernetOutput, connect to the memory allocator component.

In addition, TINET+TECS utilizes the TECS {\it send/receive} specifier to minimize the memory copy frequency, which is a functionality supported by TINET.

\begin{figure}[t]
    \centering
    \includegraphics[width=7.0cm,clip]{figure/tMemoryAllocator.pdf}
    \caption{Component diagram of tMemoryAllocator}
    \label{fig:tMemoryAllocator}
\end{figure}

\begin{figure}[t]
    \centering
    \includegraphics[width=7.0cm,clip]{figure/SendReceive.pdf}
    \caption{Differences between in/out and send/receive}
    \label{fig:SendReceive}
\end{figure}

{\bf Send/receive specifiers:}
TECS supports {\it send}/{\it receive} interface specifiers \cite{par:RPC}.
TINET+TECS uses {\it send} and {\it receive} specifiers instead of {\it in} and {\it out} to reduce the number of copies:

\begin{itemize}
\item {\it in} is a specifier for input arguments.
A callee side uses the memory of arguments with {\it in} when executing the callee function.
When the processing returns to the caller side, the caller can reuse and deallocate the memory.

\item {\it send} is another specifier for transferring data to a callee from a caller such as {\it in}.
The difference between {\it in} and {\it send} is whether the data memory is deallocated in the caller or callee, as shown in Fig.\ref{fig:SendReceive}.
In the case of the {\it in} specifier, both allocating and deallocating of the data memory are performed in the caller.
By contrast, in the case of {\it send}, the caller allocates the data memory and the callee deallocates it.

\item {\it out} is a specifier for output arguments through which a callee writes data in the memory allocated by a caller while the caller receives the data.

\item {\it receive} is another specifier for a caller receiving data from a callee such as {\it out}.
The difference between {\it out} and {\it receive} lies in whether the data memory is allocated in the caller or callee, as shown in Fig.\ref{fig:SendReceive}.
In the case of {\it out}, the callee writes data in the memory allocated by a caller, whereas in the {\it receive} case, the callee allocates the data memory.
Deallocating of the memory is performed in the caller in both cases.
\end{itemize}

As shown in Fig.\ref{src:SendReceive}, sending and receiving arguments such as {\it outputp} and {\it inputp} are defined using, respectively, the {\it send/receive} specifier in the signature description.
Developers hardly make mistakes of memory operation because these specifiers completely pass an ownership of memory.
Common object request broker architecture (CORBA) does not consider memory sharing; CORBA has no functionalities
such as {\it send}/{\it receive}.

\begin{figure}[t]
\centering
\begin{lstlisting}
signature sNicDriver {
  void start(
    [send(sNetworkAlloc),size_is(size)]
        int8_t *outputp, .., ..);
  void read(
    [receive(sNetworkAlloc),size_is(*size)]
        int8_t **inputp, .., ..);
    /* Omit: other functions */
};
\end{lstlisting}
\caption{Signature description of the nic driver (An example of send/receive)}
\label{src:SendReceive}
\end{figure}

@@

\subsubsection{Component Design of TLSF+TECS}
